The Impact of Alternation

### 2017-08-16 arxiv/draft.tex
added arxiv

In the following let $\A_i =
\tuple{\vec{x},Q_i,\iota_i,F_i,\Delta_i}$, for $i=1,2$, where
w.l.o.g. we assume that $Q_1 \cap Q_2 = \emptyset$. We define:
\[\begin{array}{rcl}
\A_\cup & = & \tuple{\vec{x},Q_1\cup Q_2,\iota_1 \vee \iota_2,F_1\cup F_2,\Delta_1\cup\Delta_2} \\ 
\A_\cap & = & \tuple{\vec{x},Q_1\cup Q_2,\iota_1 \wedge \iota_2, F_1\cup F_2, \Delta_1\cup\Delta_2} \\ 
\overline{\A_1} & = & \langle \vec{x},Q_1,\iota_1,Q_1\setminus F_1,\overline{\Delta_1} \rangle
\end{array}\] 
where $\overline{\Delta_1}(q,a) \equiv \overline{\Delta_1(q,a)}$, for
all $q \in Q_1$ and $a \in \Sigma$. The following lemma shows the
correctness of the above definitions:

@@

\vspace*{\baselineskip}
\proof{ We prove $L(\A_\cup) = L(\A_1) \cup L(\A_2)$ first, the proof
  for $\A_\cap$ being analogous. Let $w = (a_1,\nu_1) \ldots
  (a_n,\nu_n)$ be a word, where $n=0$ corresponds to the empty
  word. We prove by induction on $n\geq0$ that $\Delta(\iota_1 \vee
  \iota_2,a_1\ldots a_n) \iffeq \Delta(\iota_1,a_1\ldots a_n) \vee
  \Delta(\iota_2,a_1\ldots a_n)$. The case $n=0$ follows from the
  definition of the initial configuration of $\A_\cup$. For the
  inductive step $n>0$, $\Delta(\iota_1 \vee \iota_2,a_1\ldots a_n)$
  is obtained from $\Delta(\iota_1 \vee \iota_2,a_1\ldots a_{n-1})$ by
  replacing each variable $q \in \fv{\Bool}{\Delta(\iota_1 \vee
    \iota_2,a_1\ldots a_{n-1})}$ with
  $\Delta(q,a_n)[\vec{x}_{n-1}/\overline{\vec{x}},\vec{x}_n/\vec{x}]$,
  denoted $\Delta^n(\Delta(\iota_1 \vee \iota_2,a_1\ldots
  a_{n-1}),a_n)$. Since by induction hypothesis, $\Delta(\iota_1 \vee
  \iota_2,a_1\ldots a_{n-1}) \iffeq \Delta(\iota_1,a_1\ldots a_{n-1})
  \vee \Delta(\iota_2,a_1\ldots a_{n-1})$, we obtain:
  \[\begin{array}{ll}
  \Delta^n(\Delta(\iota_1 \vee \iota_2,a_1\ldots a_{n-1}),a_n) & \iffeq \\
  \Delta^n(\Delta(\iota_1,a_1\ldots a_{n-1}),a_n) \vee \Delta^n(\Delta(\iota_2,a_1\ldots a_{n-1}),a_n) & \iffeq \\
  \Delta(\iota_1,a_1\ldots a_{n}) \vee \Delta(\iota_2,a_1\ldots a_{n}) \enspace.
  \end{array}\] 
  To prove $L(\overline{\A_1}) = \Sigma[\vec{x}]^* \setminus L(\A_1)$,
  let $w = (a_1,\nu_1) \ldots (a_n,\nu_n)$ be a word and show that
  $\overline{\Delta}(\iota_1,a_1 \ldots a_n) =
  \overline{\Delta(\iota_1,a_1 \ldots a_n)}$ by induction on
  $n\geq0$. The case $n=0$ is immediate, because $\fv{}{\iota_1}
  \subseteq Q$ and thus $\overline{\iota_1} \equiv \iota_1$. For the case
  $n>0$, we compute:
  \[\begin{array}{lcl}
  \overline{\Delta}^n(\overline{\Delta}(\iota_1,a_1\ldots a_{n-1}),a_n) & \iffeq & \text{ (by ind. hyp.)} \\
  \overline{\Delta}^n(\overline{\Delta(\iota_1,a_1\ldots a_{n-1})},a_n) & \iffeq & \text{ (by the def. of $\overline{\phi}$)} \\
  \overline{\Delta(\iota_1,a_1\ldots,a_n)} \enspace.
  \end{array}\]
  Let $\nu,\nu' : (Q \cup \bigcup_{i=0}^n\vec{x}_i) \rightarrow (\booli
  \cup \Data^\I)$ be valuations such that: \begin{compactitem}
  \item $\nu(q) = \top$ and $\nu'(q) = \bot$, for each $q \in F$, 
    %
  \item $\nu(q) = \bot$ and $\nu'(q) = \top$, for each $q \in Q \setminus F$, 
    %
  \item $\nu(x) = \nu'(x)$, for each $x \in \vec{x}_0$, 
    %
  \item $\nu(x) = \nu'(x) = \nu_i(x)$, for each $x \in \vec{x}_i$ and each $i
    \in [1,n]$.
  \end{compactitem}
  By Proposition \ref{prop:overline}, we have $\I,\nu \models
  \Delta(\iota_1,a_1 \ldots a_n) \iff \I,\nu' \not\models
  \overline{\Delta(\iota_1,a_1 \ldots a_n)} \iff \I,\nu' \not\models
  \overline{\Delta}(\iota_1,a_1 \ldots a_n)$. Thus for all $w \in
  \Sigma[\vec{x}]^*$, we have $w \in L(\A_1) \iff w \not\in
  L(\overline{\A_1})$. \qed}

>>>>

In the following let $\A_i =
\tuple{\vec{x},Q_i,\iota_i,F_i,\Delta_i}$, for $i=1,2$, where
w.l.o.g. we assume that $Q_1 \cap Q_2 = \emptyset$. We define:
\[\begin{array}{rcl}
\A_\cup & = & \tuple{\vec{x},Q_1\cup Q_2,\iota_1 \vee \iota_2,F_1\cup F_2,\Delta_1\cup\Delta_2} \\ 
\A_\cap & = & \tuple{\vec{x},Q_1\cup Q_2,\iota_1 \wedge \iota_2, F_1\cup F_2, \Delta_1\cup\Delta_2} \\ 
\overline{\A_1} & = & \langle \vec{x},Q_1,\overline{\iota_1},Q_1\setminus F_1,\overline{\Delta_1} \rangle
\end{array}\] 
where $\overline{\Delta_1}(q,a) \equiv \overline{\Delta_1(q,a)}$, for
all $q \in Q_1$ and $a \in \Sigma$. The following lemma shows the
correctness of the above definitions:

@@

\vspace*{\baselineskip}
\proof{ We prove $L(\A_\cup) = L(\A_1) \cup L(\A_2)$ first, the proof
  for $\A_\cap$ being analogous. Let $w = (a_1,\nu_1) \ldots
  (a_n,\nu_n)$ be a word, where $n=0$ corresponds to the empty
  word. We prove by induction on $n\geq0$ that $\Delta(\iota_1 \vee
  \iota_2,a_1\ldots a_n) \iffeq \Delta(\iota_1,a_1\ldots a_n) \vee
  \Delta(\iota_2,a_1\ldots a_n)$. The case $n=0$ follows from the
  definition of the initial configuration of $\A_\cup$. For the
  inductive step $n>0$, $\Delta(\iota_1 \vee \iota_2,a_1\ldots a_n)$
  is obtained from $\Delta(\iota_1 \vee \iota_2,a_1\ldots a_{n-1})$ by
  replacing each variable $q \in \fv{\Bool}{\Delta(\iota_1 \vee
    \iota_2,a_1\ldots a_{n-1})}$ with
  $\Delta(q,a_n)[\vec{x}_{n-1}/\overline{\vec{x}},\vec{x}_n/\vec{x}]$,
  denoted $\Delta^n(\Delta(\iota_1 \vee \iota_2,a_1\ldots
  a_{n-1}),a_n)$. Since by induction hypothesis, $\Delta(\iota_1 \vee
  \iota_2,a_1\ldots a_{n-1}) \iffeq \Delta(\iota_1,a_1\ldots a_{n-1})
  \vee \Delta(\iota_2,a_1\ldots a_{n-1})$, we obtain:
  \[\begin{array}{ll}
  \Delta^n(\Delta(\iota_1 \vee \iota_2,a_1\ldots a_{n-1}),a_n) & \iffeq \\
  \Delta^n(\Delta(\iota_1,a_1\ldots a_{n-1}),a_n) \vee \Delta^n(\Delta(\iota_2,a_1\ldots a_{n-1}),a_n) & \iffeq \\
  \Delta(\iota_1,a_1\ldots a_{n}) \vee \Delta(\iota_2,a_1\ldots a_{n}) \enspace.
  \end{array}\] 
  To prove $L(\overline{\A_1}) = \Sigma[\vec{x}]^* \setminus L(\A_1)$,
  let $w = (a_1,\nu_1) \ldots (a_n,\nu_n)$ be a word and show that
  $\overline{\Delta}(\overline{\iota_1},a_1 \ldots a_n) =
  \overline{\Delta(\iota_1,a_1 \ldots a_n)}$ by induction on $n\geq0$.

  In the case $n=0$, we have $\overline{\Delta}(\overline{\iota_1},a_1
  \ldots a_n) \equiv \overline{\iota_1}$. Then $\varepsilon$ is
  accepted by $\A_1$ iff $\nu_0 \models \iota_1$, where $\nu_0(q) =
  \top$ if $q \in F_1$ and $\nu_0(q) = \bot$, otherwise. But $\nu_0
  \models \iota_1$ iff $\overline{\nu_0} \models \overline{\iota_1}$,
  where $\overline{\nu_0}(q) = \top$ if $q \not\in F_1$ and
  $\overline{\nu_0}(q) = \bot$, otherwise. Thus $\varepsilon$ is
  accepted by $\A_1$ iff it is not accepted by $\overline{\A_1}$.

  For the case $n>0$, we compute:
  \[\begin{array}{lcl}
  \overline{\Delta}^n(\overline{\Delta}(\iota_1,a_1\ldots a_{n-1}),a_n) & \iffeq & \text{ (by ind. hyp.)} \\
  \overline{\Delta}^n(\overline{\Delta(\iota_1,a_1\ldots a_{n-1})},a_n) & \iffeq & \text{ (by the def. of $\overline{\phi}$)} \\
  \overline{\Delta(\iota_1,a_1\ldots,a_n)} \enspace.
  \end{array}\]
  Let $\nu,\nu' : (Q \cup \bigcup_{i=0}^n\vec{x}_i) \rightarrow (\booli
  \cup \Data^\I)$ be valuations such that: \begin{compactitem}
  \item $\nu(q) = \top$ and $\nu'(q) = \bot$, for each $q \in F$, 
    %
  \item $\nu(q) = \bot$ and $\nu'(q) = \top$, for each $q \in Q \setminus F$, 
    %
  \item $\nu(x) = \nu'(x)$, for each $x \in \vec{x}_0$, 
    %
  \item $\nu(x) = \nu'(x) = \nu_i(x)$, for each $x \in \vec{x}_i$ and each $i
    \in [1,n]$.
  \end{compactitem}
  By Proposition \ref{prop:overline}, we have $\I,\nu \models
  \Delta(\iota_1,a_1 \ldots a_n) \iff \I,\nu' \not\models
  \overline{\Delta(\iota_1,a_1 \ldots a_n)} \iff \I,\nu' \not\models
  \overline{\Delta}(\iota_1,a_1 \ldots a_n)$. Thus for all $w \in
  \Sigma[\vec{x}]^*$, we have $w \in L(\A_1) \iff w \not\in
  L(\overline{\A_1})$. \qed}
